The zoon R package for reproducible and shareable species distribution modelling

### 2016-12-24 ms.Rmd
reworking general package info

## Lists, Chains and Replicates

@@

```{r workflow2, eval = TRUE, echo = TRUE, fig.height = 4, fig.width = 9, dpi = 300}
mosquito2 <- workflow(occurrence = UKAnophelesPlumbeus,
                      covariate  = UKBioclim,
                      process    = Chain(Background(n = 500),
                                         StandardiseCov),
                      model      = list(MaxEnt,
                                        GBM,
                                        RandomForest),
                      output     = Appify)
```

## Inspecting, sharing and extending a zoon workflow

A `zoonWorkflow` object
Structure of workflow objects:

* code (call and modules used)
* output of each module; data, results and intermediate steps
* recording the session info and package and module versions

zoon workflows are re-runnable records of the data, code and results of an entire SDM analysis and can be easily reproduced, scrutinized and extended by the whole research community.
A zoon workflow returns an object of class `zoonWorkflow`.
This object contains all the information needed to reproduce the analysis.
This includes the original call, a record of the packages and modules used and their versions when the analysis was run as well as all the data and fitted models.

A workflow is reproducible, reusable and alterable by way of the functions `RerunWorkflow` and `ChangeWorkflow`.
Using `RerunWorkflow` a workflow object can be run from scratch or from part way though the analysis to avoid computationally slow steps.
`ChangeWorkflow` takes as it's arguments a workflow object and new selections for any or all of the workflow steps.
The workflow is then rerun from the first altered step.

Things you can do to workflows:

* visualise the structure
* execute whole thing from scratch (grabs new data from web)
* execute from part way through

zoon provides functionality to quickly update an existing workflow object (even one created by someone else) by switching out modules.

Include a figure visualising the structure of the workflow object and how it can be adopted in the ChangeWorkflow function (corresponding to an example below?).

@@

#### Additional features

##### Accessing module documentation

<!-- to do -->

All modules, created by the zoon authors or contributed by others, have documentation explaining their methods and how they are used. This documentation is easily accessed in R by using the function `ModuleHelp`. This function fetched the documentation from the module repository.

##### Cross-validation

@@

##### Sharing workflows

@@

#### Example 3. What makes MaxEnt effective

@@

```{r maxnet_plots, cache = FALSE, echo = FALSE, out.width = "450px", fig.cap = "Prediction of the distribution of the Carolina wren from 5 different models, produced by the workflow `maxnet\\_comparison'. A) MaxEnt without threshold features. B) MaxNet with default settings. C) MaxNet without regularisation. D) MaxNet with only linear features. E) MaxNet without regularisation and only linear features. F) Logistic regression. MaxNet with full features but no regularisation (C) gave the most local complexity, indicative of overfitting to the data. The models with only linear features (D-F) had a broader distribution, indicative of underfitting. Differences between MaxNet with linear features and no  regularisation (E) and logistic regression (F) are due to the simple downweighting applied to data in the former."}
knitr::include_graphics("figs/maxnet_plotting-1.png")
```

>>>>

## Chains, Lists and Replicates

@@

```{r workflow2, eval = FALSE, echo = TRUE}
mosquito2 <- workflow(occurrence = UKAnophelesPlumbeus,
                      covariate  = UKBioclim,
                      process    = Chain(Background(n = 500),
                                         StandardiseCov),
                      model      = list(MaxEnt,
                                        GBM,
                                        RandomForest),
                      output     = Appify)
```

### Sharing workflows

The `zoonWorkflow` object returned by `workflow` contains all of the data, code and results used in each stage of the analysis.
The object is therefore a self-contained representation of an entire analysis which can be saved as a binary `RData` object, archived, shared with colleagues or uploaded to the web.
Anyone else may then load the object into their R session, load the zoon R package, and investigate the analysis.

### Exploring workflows

`zoonWorkflow` objects are simply R lists containing the outputs from each module in the workflow, as well as the modules used, their versions and arguments and information about the R session in which the workflow was run.

The overall structure of a `zoonWorkflow` object can be inspected and visualised using the provided `print` and `plot` methods.
The outputs and intermediate steps of the analysis can be extracted directly from the `zoonWorkflow` object, or by using the utility functions `Occurrence`, `Covariate`, `Process`, `Model` and `Output` to extract the outputs of each analytical stage.
These functions return either the single R object outputted by each module, or list of objects when a list of modules was used in the workflow.

### Reproducing and extending workflows

Given a `zoonWorkflow` object, the analysis can be repeated in its entirety using the zoon function `RerunWorkflow`.
This function can therefore be used to rapidly update an analysis whenever the underlying occurrence or covariate data aree updated.
`RerunWorkflow` can also run from a specific stage of the workflow, for example using the previously downloaded data stored in the object but re-running an output module to generate a plot.

Workflows can similarly be modified and re-run by replacing or adding modules in one or more of the five analytical steps using the function `ChangeWorkflow`.
`ChangeWorkflow` only re-runs the channged modules and those downstream; the stored outputs of earlier modules are re-used rather than re-running them.

This functionality therefore enables researchers to explore and alter existing workflows, such as from a published analysis, without having to re-run computationally expensive models or re-download datasets that may have changed in the interim. See example 2 below for a demonstration of this functionality.

### Exploring available modules

zoon modules are not distributed with the zoon R packages, but are instead downloaded on-the-fly for the online module repository.
New modules uploaded to the repository therefore become instantly available to zoon users, without having to update the R package.
zoon can query the currently available modules of each type via the function `GetModuleList`.

All zoon modules on the repository are accompanied by documentation and metadata, similarly to R's helpfiles.
The documentation for any module on the repository can be accessed from R using the function `ModuleHelp`.
zoon also provides a function `ZoonCitation` to provide citation information; similarly to the `citation` function provided with R for citing R packages.
A web service for interactive exploration of available modules and their documentation will also be made available soon.

@@

### Cross-validation

@@

@@

#### Example 3. Unpacking MaxEnt

@@

```{r maxnet_plots, cache = FALSE, echo = FALSE, out.width = "450px", fig.cap = "Prediction of the distribution of the Carolina wren from 5 different models, produced by the workflow `maxnet\\_comparison'. A) MaxEnt without threshold features. B) MaxNet with default settings. C) MaxNet without regularisation. D) MaxNet with only linear features. E) MaxNet without regularisation and only linear features. F) Logistic regression. MaxNet with full features but no regularisation (C) gave the most local complexity, indicative of overfitting to the data. The models with only linear features (D-F) had a broader distribution, indicative of underfitting. Differences between MaxNet with linear features and no  regularisation (E) and logistic regression (F) are due to the downweighting applied to background data in the former."}
knitr::include_graphics("figs/maxnet_plotting-1.png")
```
